---
title: "Hemphill R Package"
author: "Sydney Hemphill"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation
Load in common libraries used by this package.

```{r eval = FALSE}
library(ggplot2)
library(tidyverse)
```

Now, we install my R package.

```{r eval = FALSE}
devtools::install_github("sydneyhemphill/R_Package_Hemphill")
```

Let us download some data for our package:
```{r}
download.file("https://raw.githubusercontent.com/sydneyhemphill/R_package_Hemphill/master/data/Butterfly_data.csv" , destfile = "/cloud/project/data/butterfly_data.csv")
```

## Working Example

Statistical Test:
This function is helpful because it takes out the chance of any mistakes while typing and will save time. The output of the data should be a summary of the linear model for your data. 
```{r} 
library(HemphillRPackage)
linear.model <- function(x,y,z){
  model.fit<- lm(x ~ y, data = z)
  return(summary(model.fit))
}
```
For an example of this function in use:
```{r}
library(HemphillRPackage)
linear.model(butterfly.data$SpringTemp,butterfly.data$Year, butterfly.data)
```

Plot Function:
This function is helpful because it creates basic plots, so you don't have to type out everything when comparing data or you need a lot of basic graphs for your reports. The output of the data should be a simple gg (scatter) plot.

```{r}
library(HemphillRPackage)
plot<- function(z, x, y){
  product<- ggplot(z, aes(x=x, y = y)) + geom_point()
  return (product)
}
```
An example of this function in use:

```{r}
library(HemphillRPackage)
plot(butterfly.data, butterfly.data$Year, butterfly.data$Day )
```

Clean Up Function:
This function removes NAs from databases. Removing NAs can take a long time, especially with large databases, so this function takes all the work, and possibility of error, out by removing the NAs as the file is read in.

```{r}
library(HemphillRPackage)
no.na<- function(file = "/cloud/project/data/Butterfly_data.csv"){
      butterfly<- read_csv(file)
      butterfly <- na.omit(butterfly)
      return(butterfly)
   }
```
An example of this function in action is:
```{r}
library(HemphillRPackage)
no.na("/cloud/project/data/Butterfly_data.csv")
```
